### 2019-11-30 report/02-methods.Rmd
structure project 1 and 2

@@

habitat_station <- hab_analysis %>%
  select(-opcod, -aqua_vg, - shade, - sinuosite) %>%
  mutate_if(is.character, list(~factor(., levels = lvl, ordered = TRUE))) %>%
  left_join(select(hab_analysis, station, shade, sinuosite)) %>%
  mutate(
    shade = factor(shade, levels = lvl_shade, ordered = TRUE),
    sinuosite = factor(sinuosite, levels = lvl_sinuosite, ordered = TRUE)) %>%
  group_by(station) %>%
  summarise_all(list(med = ~median(as.numeric(.), na.rm = TRUE))) %>%
  left_join(select(temporal_station_desc, station, width_river_mean, avg_depth_station_mean))

# Add alt, lat, etc...
habitat_station %<>% 
  left_join(geo_station, by = "station") %>%
  select(-region_csp)
# Add temp and flow
flow_temp <- temporal_press_polluants %>%
  filter(category %in% c("flow", "temperature", "DBO", "herbicides",
      "fungicides", "insecticides", "nitrates", "phosphore")) %>%
  rename(station = id) %>%
  select(station, category, press_med) %>%
  spread(category, press_med)

habitat_station %<>%
  left_join(flow_temp, by = "station")
habitat_pressure <- habitat_station
mysave(habitat_pressure, dir = mypath("report"), overwrite = TRUE)
```

@@

```{r pca-hab, fig.dim = c(10, 7), fig.cap = "PCA over habitat description"}
library(ade4)
library(factoextra)
pca_hab <- dudi.pca(as.data.frame(select(na.omit(habitat_station), -station)), scannf = FALSE, nf = 2, center = TRUE, scale = TRUE)

eig_plot <- fviz_eig(pca_hab)
pca_plot <- fviz_pca_var(pca_hab,
    axes = c(1, 2),
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE     # Avoid text overlapping
  )

ggdraw(pca_plot) + 
  draw_plot(eig_plot, x = .7, y = .95, hjust = 1, vjust = 1, width = 0.17, height = 0.25)
```

@@

@@

dist_env <- habitat_pressure %>%
  select(station, width_river_mean:temperature) %>%
  gather(env_var, value, width_river_mean:temperature) 
med_dist_env <- dist_env %>%
  group_by(env_var) %>%
  summarise(value = median(value, na.rm = TRUE))

@@

@@

```{r}
library(vegan)

@@

```{r}
# Add habitat
match_st <- match(habitat_pressure$station, com$station)
habitat <- habitat_pressure[order(match_st), ]
match_st_com <- match(temporal_community_metrics$station, com$station)
com_metrics <- temporal_community_metrics[order(match_st_com), ]

ordisurf(nmds,habitat$alt,main="",col="forestgreen")
```

```{r}

hab_var <- c("temperature", "alt", "lat", "flow", "width_river_mean", "DBO")
par(mfrow = c(2,3))
for (i in seq_along(hab_var)) {
  ordisurf(nmds, habitat[[hab_var[i]]], main = hab_var[i], col = "brown")
}

@@

```{r basin-com}
match_st_basin <- match(station_analysis$station, com$station)
st_basin <- station_analysis[order(match_st_basin), ]

ordiplot(nmds,type="n")
orditorp(nmds,display="species",col="red",air=0.01)
ordihull(nmds, groups = st_basin[["basin"]], draw="polygon",col="grey90",label=F)
```

@@

$$
C_{k} = CV_{-k} - CV
$$

@@

```{r load dataset, include = FALSE}
library(jmotif)
myload(community_metrics, op_analysis, dir = mypath("data"))
community_metrics %<>% left_join(select(op_analysis, opcod, station, date))
```

>>>>

### Fish length

```{r, fig.cap = "Distribution of fish length. Red, green and black respectively locate the median, the mean and the 200mm threshold."}
myload(length_analysis, dir = mypath("data"))
length_analysis %>%
  ggplot(aes(x = length)) +
  geom_histogram() +
  geom_vline(xintercept = median(length_analysis$length, na.rm = TRUE), color = "red") +
  geom_vline(xintercept = mean(length_analysis$length, na.rm = TRUE), color = "green") +
  geom_vline(xintercept = 200, color = "black")
```

@@

habitat_station <- hab_analysis %>%
  select(-opcod, -aqua_vg, - shade, - sinuosite) %>%
  mutate_if(is.character, list(~factor(., levels = lvl, ordered = TRUE))) %>%
  left_join(select(hab_analysis, station, shade, sinuosite)) %>%
  mutate(
    shade = factor(shade, levels = lvl_shade, ordered = TRUE),
    sinuosite = factor(sinuosite, levels = lvl_sinuosite, ordered = TRUE)) %>%
  group_by(station) %>%
  summarise_all(list(med = ~median(as.numeric(.), na.rm = TRUE))) %>%
  left_join(select(temporal_station_desc, station, width_river_mean, avg_depth_station_mean, width_river_cv, avg_depth_station_cv))

# Add alt, lat, etc...
habitat_station %<>% 
  left_join(geo_station, by = "station") %>%
  select(-region_csp)
# Add temp and flow
flow_temp <- temporal_press_polluants %>%
  filter(category %in% c("flow", "temperature", "DBO", "herbicides",
      "fungicides", "insecticides", "nitrates", "phosphore")) %>%
  rename(station = id) 
flow_temp_med <- flow_temp %>%
  mutate(category = str_c(category, "_med")) %>%
  select(station, category, press_med) %>%
  spread(category, press_med)
flow_temp_cv <- flow_temp %>%
  mutate(category = str_c(category, "_cv")) %>%
  select(station, category, cv_press) %>%
  spread(category, cv_press)

habitat_station %<>%
  left_join(left_join(flow_temp_med, flow_temp_cv, by = "station"), by = "station")
habitat_pressure <- habitat_station
mysave(habitat_pressure, dir = mypath("report"), overwrite = TRUE)
```

@@

```{r pca-hab, fig.dim = c(7, 21), fig.cap = "PCA over habitat description"}
library(ade4)
library(factoextra)
pca_hab <- dudi.pca(as.data.frame(select(na.omit(habitat_station), -station)), scannf = FALSE, nf = 3, center = TRUE, scale = TRUE)

eig_plot <- fviz_eig(pca_hab)
pca_plot <- fviz_pca_var(pca_hab,
    axes = c(1, 2),
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE     # Avoid text overlapping
  )
pca_plot12 <- ggdraw(pca_plot) +
  draw_plot(eig_plot, x = .7, y = .95, hjust = 1, vjust = 1, width = 0.17, height = 0.25)

pca_plot13 <- fviz_pca_var(pca_hab,
    axes = c(1, 3),
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE     # Avoid text overlapping
  )

pca_plot23 <- fviz_pca_var(pca_hab,
    axes = c(2, 3),
    col.var = "contrib", # Color by contributions to the PC
    gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
    repel = TRUE     # Avoid text overlapping
  )
plot_grid(pca_plot12, pca_plot13, pca_plot23, ncol = 1)
```

@@

```{r}
dev.off()
m <- cor(as.matrix(select(na.omit(habitat_station), -station)), method = "spearman")
corrplot::corrplot(m, order="hclust", type = "upper", diag = FALSE)
```

@@

dist_env <- habitat_pressure %>%
  select(station, width_river_mean:temperature_cv) %>%
  gather(env_var, value, width_river_mean:temperature_cv) 
med_dist_env <- dist_env %>%
  group_by(env_var) %>%
  summarise(value = median(value, na.rm = TRUE))

@@

```{r split-rivers, fig.dim = c(7, 12)}
hab_type <- habitat_pressure %>%
  mutate(high_bin = ifelse(alt > median(alt), TRUE, FALSE),
  wide_bin = ifelse(width_river_mean > median(width_river_mean), TRUE, FALSE),
  hab_type = map2_chr(high_bin, wide_bin, function(high, wide){
    if (high) {
      if (wide) {
	"high_wide"
      } else {
	"high_narrow"
      }
    } else {
      if (wide) {
	"low_wide"
      } else {
	"low_narrow"
      }
    }

      })
  ) %>%
mutate(hab_type = map_chr(high_bin, ~ifelse(.x, "high", "low"))) %>%
select(station, hab_type)

hab_type_pca <- hab_type[hab_type$station %in% na.omit(habitat_station)$station, ]
match_st <- match(hab_type_pca$station, na.omit(habitat_station)$station)
hab_type_pca <- hab_type_pca[order(match_st), ]

pca_cat_river <- fviz_pca_ind(pca_hab, label="none", habillage=hab_type_pca$hab_type,
  , ellipse.type = "convex",
             addEllipses=TRUE, ellipse.level=0.95)
plot_grid(pca_plot, pca_cat_river, ncol = 1)
mysave(hab_type, dir = mypath("report"), overwrite = TRUE)
```

@@

```{r, echo = FALSE}
library(vegan)

@@

```{r}
# Add habitat
match_st <- match(habitat_pressure$station, com$station)
habitat <- habitat_pressure[order(match_st), ]
match_st_com <- match(temporal_community_metrics$station, com$station)
com_metrics <- temporal_community_metrics[order(match_st_com), ]
```

```{r, fig.dim = c(7, 14)}
hab_var <- c("temperature_med", "temperature_cv", "alt", "lat", "flow_med", "flow_cv", "width_river_mean", "width_river_cv", "DBO_med", "DBO_cv")
par(mfrow = c(5,2))
for (i in seq_along(hab_var)) {
  ordisurf(nmds, habitat[[hab_var[i]]], main = hab_var[i], col = "brown")
}
```

@@

```{r basin-com}
dev.off()
match_st_basin <- match(station_analysis$station, com$station)
st_basin <- station_analysis[order(match_st_basin), ] %>%
  mutate(basin = ifelse(is.na(basin), "INCONNU", basin))

ordiplot(nmds,type="n")
orditorp(nmds,display="species",col="red",air=0.01)
# See https://www.davidzeleny.net/anadat-r/doku.php/en:ordiagrams_examples
u_basin  <- unique(st_basin[["basin"]])
for (i in seq_along(u_basin)) {
  ordihull(nmds, groups = st_basin[["basin"]],
 show.group = u_basin[i], col = i, draw = 'polygon', label = T)
}
```

```{r}
match_st <- match(hab_type$station, com$station)
hab_type <- hab_type[order(match_st), ]
ordiplot(nmds,type="points", choices = c(1, 2))
orditorp(nmds,display="species",col="red",air=0.01)
u_hab  <- unique(hab_type[["hab_type"]])
for (i in seq_along(u_hab)) {
  ordihull(nmds, groups = hab_type[["hab_type"]], alpha = .1,
 show.group = u_hab[i], col = i, draw = 'polygon', label = T)
}
```

```{r}
st_basin <- get_basin_station() 
myload(region_polygon, dir = mypath("data"))
myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
stab_div_basin <- temporal_network_metrics_classes %>%
  select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group)) %>%
  select(station, troph_group, biomass_stab, richness_avg) %>%
  left_join(st_basin, by = "station") %>%
  filter(!is.na(basin))
stab_div_basin %>%
  left_join(hab_type, by = "station") %>%
  left_join(habitat_pressure[,c("station", "temperature_med", "DBO_med")],
    by = "station") %>%
  gather(variable, value, temperature_med, DBO_med) %>%
  ggplot(aes(x = richness_avg, y = value)) +
  geom_point() +
  facet_grid(cols = vars(troph_group), rows = vars(variable), scales = "free_y") +
  stat_poly_eq(
    formula = y ~ x,
    eq.with.lhs = "italic(hat(y))~`=`~",
    aes(label = paste(..eq.label.., ..rr.label.., sep = "*plain(\",\")~")),
    parse = TRUE) +
  geom_smooth(method = "lm", formula = y ~ x) +
  ggplot2::scale_color_manual(
    labels = troph_group_labeller(),
    values = c("blue", "green", "red"),
    name = "Trophic level"
    ) +
  #scale_y_log10() +
  scale_x_log10() +
  labs(title = "Stability - Div by habitat type",
    x = "Average species richness (log10)", y = "Biomass stability (log10)")
```

@@

$$
C_{k} = X - X_{-k}
$$

It means that a negative $C_k$ correspond to a negative contribution of the
species $k$ to $X$.

@@

```{r load dataset}
library(jmotif)
myload(community_metrics, op_analysis, dir = mypath("data"))
community_metrics  <-   
  left_join(community_metrics, op_analysis[, c("opcod", "station", "date")], by = "opcod")
```
